---
title: "p8105_hw2_hm2900"
output: github_document
---

```{r setup}
library(tidyverse)
```

# Problem 1

Read and clean the data.retain line, station, name, station latitude / longitude, routes served, entry, vending, entrance type, and ADA compliance.

```{r}
nyc_transit_df = 
  read_csv(
    "./NYC_Transit_Subway_Entrance_And_Exit_Data.csv", 
    col_types = cols(Route8 = "c", Route9 = "c", Route10 = "c", Route11 = "c")) %>% 
  janitor::clean_names() %>%
  select(
    line, station_name, station_latitude, station_longitude, 
    starts_with("route"), entry, exit_only, vending, entrance_type, 
    ada) %>% 
  mutate(entry = recode(entry, "YES" = TRUE, "NO" = FALSE))
```

The data set contains 19 variables: `line`, `station_name`, ``station_latitude`, `station_longitude`, `route1`-`route11`, `entry`, `exit_only`, `entrance_type`, `vending`, and `ada`. 
For the data cleaning steps, first I cleaned the variables' names in the dataset, then I used selected the columns that I want. Finally I changed the entry variable from character (YES vs NO) to a logical variable. The dimension of the resulting dataset is 1,868 × 19. 
These data are not tidy enough since route number should be a variable, as should route. That is, to obtain a tidy dataset we would need to convert `route` variables from wide to long format.

## How many distinct stations are there?

```{r}
nyc_transit_df %>% 
  select(station_name, line) %>% 
  distinct
```

There are 465 distinct stations.

## How many stations are ADA compliant?

```{r}
nyc_transit_df %>% 
  filter(ada == TRUE) %>%
  select(station_name, line) %>% 
  distinct 
```

There are 84 stations that are ADA compliant.

## What proportion of station entrances / exits without vending allow entrance?

```{r}
nyc_transit_df %>%
  filter(vending == "NO") %>% 
  pull(entry) %>% 
  mean
```

The proportion station entrances / exits without vending is 37.7%.

## How many distinct stations serve the A train? Of the stations that serve the A train, how many are ADA compliant?

```{r}
nyc_transit_df %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_number",
    values_to = "route") %>% 
  filter(route == "A") %>% 
  select(station_name, line) %>% 
  distinct

nyc_transit_df %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_number",
    values_to = "route") %>% 
  filter(route == "A", ada == TRUE) %>% 
  select(station_name, line) %>% 
  distinct
```

There are 60 distinct stations serve the A train. Of the stations that serve the A train, there are 17 stations that are ADA compliant.












































